### 2022-10-16 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@e411d6fd2bc3a08964824a43ec81f379ae7c2be2 🚀

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes. Still
a robust understanding of, and ability to predict and forecast
interactions among species and their consequences for network structure
remains elusive. This knowledge-gap is largely driven by a shortfall of
data---although species occurrence data has rapidly increased in the
last decade, species interaction data has not kept pace, largely due to
the intrinsic difficulty and effort required to sample interactions.
These sampling challenges bias data and hinder inferences about the
structure and dynamics of interactions networks. Here, we demonstrate
the realized false-negative rate (the percentage of species that
actually interact but for which we do not yet have a record) can be
quite high, even in thoroughly sampled systems, due to the intrinsic
variation in abundances across species in a community. We demonstrate
how a null model of occurrence detection can be used to estimate the
false-negative rate in a given dataset. One hypothesis is that
interactions between ``rare'' species are themselves rare because these
species are less likely to encounter one-another than species of higher
relative abundance. However, we demonstrate that across several datasets
of spatial or temporally replicated networks, there are positive
associations that suggest these interactions actually exist but just are
not observed. Finally, we assess how false negatives influence various
models of network prediction, and recommend directly accounting for
observation error in predictive models. We conclude by discussing how
the understanding of false-negatives can inform how we design monitoring
schemes for species interactions.\\%

@@

The importance of \emph{sampling effort} and its impact on resulting
ecological data has produced a rich body of literature. The recorded
number of species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001)---as do
estimates of population abundance (Griffiths 1998)---in addition to
spatial coverage and species detectability. This has motivated more
quantitatively robust approaches to account for error in sampling data
in many contexts: to determine if a given species is extinct (Boakes
\emph{et al.} 2015), to determine sampling design (Moore \& McCarthy
2016),and to measure species richness across large scales (Carlson
\emph{et al.} 2020). In the context of interactions, an initial concern
was the compounding effects of limited sampling effort combined with the
amalgamation of data (across both study sites, time of year, and
taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact (Paine 1988) or
the structure of the network as a whole (McLeod \emph{et al.} 2021).
Martinez \emph{et al.} (1999) showed that in a plant-endophyte trophic
network, network connectance is robust to sampling effort, but this was
done in the context of a system for which observation of 62,000 total
interactions derived from 164,000 plant-stems was feasible. In some
systems (e.g.~megafauna food-webs) this many observations is either
impractical or infeasible due to the absolute abundance of the species
in question.

We cannot feasibly observe all (or even most) of the direct interactions
that occur in an ecosystem. This means we can be confident two species
actually interact if we have a record of it (given an estimate of
species misidentification probability), but not at all confident that a
pair of species \emph{do not} interact if we have no record of those
species observed together. In other words, it is difficult to
distinguish \emph{true-negatives} (two species never interact) from
\emph{false-negatives} (two species interact sometimes, but we do not
have a record of it). For a concrete example of a false-negative in a
food web, see fig.~\ref{fig:concept}. Because even the most highly
sampled systems will still contain missing interactions, there is
increasing interest in combining species-level data (e.g.~traits,
abundance, range, phylogenetic relatedness, etc.) to build models to
predict interactions between species we haven't observed together before
(Strydom \emph{et al.} 2021). However, the noise of false-negatives
could impact the efficacy of our predictive models and have practical
consequences for answering questions about interactions (de Aguiar
\emph{et al.} 2019). This data constraint is amplified as the
interaction data we have is geographically biased toward the usual
suspects (Poisot \emph{et al.} 2021). We therefore need a systematic
approach to assessing these biases in the observation process and the
consequences this has for our understanding of interaction networks.

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another directly in proportion to each
species' relative-abundance. This assumes that there are no associations
in species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are rampant in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case (Blanchet \emph{et al.} 2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, exhibit any meaningful effect on the fitness of the other. So,
although co-occurrence is not directly indicative of an interaction, it
is a precondition for an interaction. Therefore observations of
``non-interactions'' between pairs of species that are outside of the
union of both species ranges do not provide any information about that
interaction, i.e.~they should be excluded from consideration.

Here, we illustrate how our confidence that a pair of species we believe
to not interact highly depends on sampling effort, and suggest that
surveys of species interactions can benefit from simulation modeling of
detection probability. We demonstrate that the realized false-negative
rate of interactions is directly related to the relative abundance of
the species involved, and demonstrate how simulation can be used to
produce a null estimate of the false-negative rate as a function of
total sampling effort (the total count of all individuals of all species
seen). We show that positive associations in co-occurrence data can
increase realized probability of false-negatives, and demonstrate these
positive associations are ubiquitous in network datasets. We conclude by
recommending that the simulation of sampling effort and species
occurrence can and should be used to help design surveys of species
interaction diversity (Moore \& McCarthy 2016), and by advocating use of
null models like those presented here as a tool for both guiding design
of surveys of species interactions and for modeling detection error in
predictive models.

@@

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as
not-interacting with an independent and fixed probability, which we
denote pfn and subsequently refer to as the False-Negative Rate (FNR).
If we observe the same species not-interacting \(N\) times, then the
probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the probability-mass-function of
geometric distribution, a special case of the negative-binomial
distribution) is shown in Figure 1(A) for varying values of \(p_{fn}\)
and illustrates a fundamental link between our ability to reliably say
an interaction doesn't exist---\(p_{tn}\)---and the number of times we
have observed a given species. In addition, note that also there is no
non-zero \(p_{fn}\) for which we can ever prove that an interaction does
not exist---no matter how many observations of non-interactions \(N\) we
have, \(p_{tn}<1\).

From fig.~\ref{fig:geometric}(A) it is clear that the more often we see
two species co-occurring, but not interacting, the more likely the
interaction is a true negative. This has several practical consequences:
first it means negatives taken outside the overlap of the range of each
species aren't informative. Second, we can use this relation to compute
the expected number of total observations needed to obtain a ``goal''
number of observations of a particular pair of species
(fig.~\ref{fig:geometric}(B)). As an example, if we hypothesize that
\(A\) and \(B\) do not interact, and we want to see species \(A\) and
\(B\) both co-occurring and not interacting 10 times to be confident
this is a true negative, then we need an expected 1000 observations of
all species if the relative abundances of \(A\) and \(B\) are both
\(0.1\).

@@

\begin{figure}
\hypertarget{fig:geometric}{%
\centering
\includegraphics{./figures/fig1.png}
\caption{(A) The probability that an observed interaction is a true
negative (y-axis) given how many times it has been sampled as a
non-interaction (x-axis). Each color reflects a different value of
\(p_{fn}\), the false-negative rate (FNR)---this is effectively the cdf
of the geometric distribution. (B): The expected needed observations of
all individuals of all species (y-axis) required to obtain a goal number
of observations (colors) of a particular species, and a function of the
relative abundance of that focal species (x-axis). (C): False negative
rate (y-axis) as a function of total sampling effort (x-axis) and
network size, computed using the method described above. For 500
independent draws from the niche model (Williams \& Martinez (2000)) at
varying levels of species richness (colors) with connectance drawn
according to the flexible-links model (MacDonald \emph{et al.} (2020))
as described in the main text. For each draw from the niche model, 200
sets of 1500 observations are simulated, for which each the mean false
negative rate at each observation-step is computed. Means denoted with
points, with 1 in the first shade and 2 in the second. (D): Same as (C),
except using empirical food webs from Mangal database, where richness.
The outlier on (D) is a 714 species food-web.}\label{fig:geometric}
}
\end{figure}

@@

We now show that the realized FNR changes drastically with sampling
effort due to the intrinsic variation of the abundance of individuals of
each species within a community. We do this by simulating the process of
observation of species interactions, applied both to 243 empirical food
webs from the Mangal database (Banville \emph{et al.} 2021) and random
food-webs generated using the niche model (Williams \& Martinez 2000), a
simple generative model of food-web structure. Our neutral model of
observation assumes each observed species is drawn from the distribution
in proportion to each species' abundance at that place and time. The
abundance distribution of a community can be reasonably-well described
by a log-normal distribution (Volkov \emph{et al.} 2003). In addition to
the log-normal distribution, we also tested the case where the abundance
distribution is derived from power-law scaling \(Z^{(log(T_i)-1)}\)
where \(T_i\) is the trophic level of species \(i\) and \(Z\) is a
scaling coefficient (Savage \emph{et al.} 2004), which yields the same
qualitative behavior. The practical consequence of abundance
distributions spanning many orders of magnitude of abundance is that
observing two ``rare'' species interacting requires two low probability
events: observing two rare species at the same time.

To simulate the process of observation, for an ecological network \(M\)
with \(S\) species, we sample abundances for each species from a
standard-log-normal distribution. For each true interaction in the
adjacency matrix \(M\) (i.e. \(M_{ij}=1\)) we estimate the probability
of observing both species \(i\) and \(j\) at a given place and time by
simulating \(n\) observations of all individuals of any species, where
the species of the individual observed at the \(\{1,2,\dots,n\}\)-th
observation is drawn from the generated log-normal distribution of
abundances. For each pair of species \((i,j)\), if both \(i\) and \(j\)
are observed within the n-observations, the interaction is tallied as a
true positive if \(M_{ij}=1\). If only one of \(i\) or \(j\) are
observed---but not both---in these \(n\) observations, but \(M_{ij}=1\),
this is counted as a false-negative, and a true-negative otherwise.

@@

where the MAP estimate of \((\mu, \phi)\) applied to Mangal data from
(MacDonald \emph{et al.} 2020) is \((\mu=0.086, \phi=24.3)\). All
simulations were done with 500 independent replicates per unique number
of observations \(n\). All analyses presented here are done in Julia
v1.8 (Bezanson \emph{et al.} 2015) using both EcologicalNetworks.jl v0.5
and Mangal.jl v0.4 (Banville \emph{et al.} 2021) and are hosted at
(GITHUB\_LINK\_TODO). Note that the empirical data, for the reasons
described above, very likely already contains many false negatives,
we'll revisit this issue in the final section.

From fig.~\ref{fig:geometric}(C) it is evident that the number of
species considered in a study is inseparable from the false-negative
rate in that study, and this effect should be taken into account when
designing samples of ecological networks in the future. We see a similar
qualitative pattern in fig.~\ref{fig:geometric}(D) where the FNR drops
off quickly as a function of observation effort, mediated by total
richness. The practical consequence of the bottom row of Figure 1 is
whether the total number of observations of all species (the x-axis) for
the range of possible FNR we deem acceptable (the y-axis) is feasible.
This raises two points: first, empirical data on interactions are
subject to the practical limitations of funding and human-work hours,
and therefore existing data tend to fall on the order of hundreds or
thousands observations of individuals per site. Clear aggregation of
data on sampling effort has proven difficult to find and a meta-analysis
of network data and sampling effort seems both pertinent and necessary,
in addition to the effects of aggregation of interactions across
taxonomic scales (Gauzens \emph{et al.} 2013; Giacomuzzo \& Jordán
2021). This inherent limitation on in-situ sampling means we should
optimize where we sample across space so that for a given number of
samples, we obtain the maximum information possible. Second, what is
meant by ``acceptable'' FNR? This raises the question: does a shifting
FNR lead to rapid transitions in our ability inference and predictions
about the structure and dynamics of networks, or does it produce a
roughly linear decay in model efficacy? We explore this in the next
section.

We conclude this section by advocating for the use of neutral models
similar to above to generate expectations about the number of
false-negatives in a data set of a given size. This could prove fruitful
both for designing surveys of interactions but also because we may want
to incorporate models of imperfect detection error into predictive
interactions models, as Joseph (2020) does for species occurrence
modeling. Additionally, we emphasize that one must consider the context
for sampling---is the goal to detect a particular species (as in
fig.~\ref{fig:geometric}(C)), or to get a representative sample of
interactions across the species pool? This argument is well-considered
when sampling individual species (Willott 2001), but has not yet been
adopted for designing samples of communities.

@@

This model above doesn't consider the possibility that there are
positive or negative associations which shift the probability of
observing two species together due to their interaction (Cazelles
\emph{et al.} 2016). However, here we demonstrate that the probability
of observing a false negative can be higher if there is some positive
association in the occurrence of species \(A\) and \(B\). If we denote
the probability that we observe the co-occurrence of two species \(A\)
and \(B\) that we know interact as \(P(AB)\) and if there is no
association between the marginal probabilities of observing \(A\) and
observing \(B\), denoted \(P(A)\) and \(P(B)\) respectively, then the
probability of observing their co-occurrence \(P(AB) = P(A)P(B)\). In
the other case where there is some positive strength of association
between observing both \(A\) and \(B\) because this interaction is
``important'' for each species, then the probability of observation both
\(A\) and \(B\), \(P(AB)\), is greater than \(P(A)P(B)\) as \(P(A)\) and
\(P(B)\) are not independent and instead are positively correlated,
i.e.~\(P(AB)> P(A)P(B)\). In this case, the probability of observing a
false negative in our naive model from fig.~\ref{fig:geometric}(A) is
\(p_{fn}= 1-P(AB)\), which due to the above inequality implies
\(p_{fn} >1-P(A)P(B)\). This indicates an increasingly greater
probability of a false negative as \(P(AB) \to P(AB) \gg P(A)P(B)\).
However, this still does not consider variation in species abundance in
space and time (Poisot \emph{et al.} 2015). If positive or negative
associations between species structure variation in the distribution of
\(P(AB)\) across space/time, then the spatial/temporal biases induced by
data collection would further impact the realized false negative rate,
as the probability of false negative would not be constant for each pair
of species across sites.

To test for this association in data we scoured Mangal for datasets with
many spatial or temporal replicates of the same system. For each
dataset, we compute the marginal probability \(P(A)\) of occurrence of
each species \(A\) across all networks in the dataset. For each pair of
interacting species \(A\) and \(B\), we then compute and compare the
probability of co-occurrence if each species occurs independently,
\(P(A)P(B)\), to the empirical joint probability of co-occurrence,
\(P(AB)\). Following our analysis above, if \(P(AB)\) is greater than
\(P(A)P(B)\), then we expect our neutral estimates of the FNR above to
underestimate the realized FNR. In fig.~\ref{fig:mangal}, we see the
difference between \(P(AB)\) and \(P(A)P(B)\) for the seven suitable
datasets with enough spatio-temporal replicates and a shared taxonomic
backbone (meaning all individual networks use common species
identifiers) found on Mangal to perform this analysis. Further details
about each dataset are reported in tbl.~\ref{tbl:id}.

\begin{figure}
\hypertarget{fig:mangal}{%
\centering
\includegraphics{./figures/fig2.png}
\caption{The difference between joint-probability of co-occurrence
(\(P(AB)\)) and expected probability of co-occurrence under independence
(\(P(A)P(B)\)) for interacting species for each dataset. The red-dashed
line indicates 0. Each histogram represents a density, meaning the area
of the entire curve sums to 1. The continuous density estimate (computed
using local smoothing) is shown in grey. The p-value on each plot is the
result of a one-sided t-test comparing the mean of each distribution to
0.}\label{fig:mangal}
}
\end{figure}

>>>>

        {\bfseries Abstract:}\,Ecosystems are composed of networks of
interacting species. These interactions allow communities of species to
persist through time through both neutral and adaptive processes. Still
a robust understanding of, and ability to predict and forecast
interactions among species remains elusive. This knowledge-gap is
largely driven by a shortfall of data---although species occurrence data
has rapidly increased in the last decade, species interaction data has
not kept pace, largely due to the intrinsic difficulty and effort
required to sample interactions. These sampling challenges bias data and
hinder inferences about the structure and dynamics of interactions
networks. Here, we demonstrate the realized false-negative rate (the
percentage of species that actually interact but for which we do not yet
have a record) can be quite high, even in thoroughly sampled systems,
due to the intrinsic variation in abundances across species in a
community. We illustrate how a null model of occurrence detection can be
used to estimate the false-negative rate in a given dataset. One
hypothesis is that interactions between ``rare'' species are themselves
rare because these species are less likely to encounter one-another than
species of higher relative abundance. However, we demonstrate that
across several datasets of spatial or temporally replicated networks,
there are positive associations that suggest these interactions actually
exist but just are not observed. Finally, we assess how false negatives
influence various models of network prediction, and recommend directly
accounting for observation error in predictive models. We conclude by
discussing how the understanding of false-negatives can inform how we
design monitoring schemes for species interactions.\\%

@@

The importance of \emph{sampling effort} and its impact on resulting
ecological data has produced a rich body of literature. The recorded
number of species in a dataset or sample depends on the total number of
observations (Walther \emph{et al.} 1995; Willott 2001)---as do
estimates of population abundance (Griffiths 1998)---in addition to
spatial coverage and species detectability. This has motivated more
quantitatively robust approaches to account for error in sampling data
in many contexts: to determine if a given species is extinct (Boakes
\emph{et al.} 2015), to determine sampling design (Moore \& McCarthy
2016), and to measure species richness across large scales (Carlson
\emph{et al.} 2020). In the context of interactions, an initial concern
was the compounding effects of limited sampling effort combined with the
amalgamation of data (across both study sites, time of year, and
taxonomic scales) could lead any empirical set of observations to
inadequately reflect the reality of how species interact (Paine 1988) or
the structure of the network as a whole (Martinez \emph{et al.} 1999;
McLeod \emph{et al.} 2021). Martinez \emph{et al.} (1999) showed that in
a plant-endophyte trophic network, network connectance is robust to
sampling effort, but this was done in the context of a system for which
observation of 62,000 total interactions derived from 164,000
plant-stems was feasible. In some systems (e.g.~megafauna food-webs)
this many observations is either impractical or infeasible due to the
absolute abundance of the species in question.

We cannot feasibly observe all (or even most) of the interactions that
occur in an ecosystem. This means we can be confident two species
actually interact if we have a record of it (given an estimate of
species misidentification probability), but not at all confident that a
pair of species \emph{do not} interact if we have \emph{no record} of
those species observed together. In other words, it is difficult to
distinguish \emph{true-negatives} (two species never interact) from
\emph{false-negatives} (two species interact sometimes, but we do not
have a record of it). For a concrete example of a false-negative in a
food web, see fig.~\ref{fig:concept}. Because even the most highly
sampled systems will still contain missing interactions, there is
increasing interest in combining species-level data (e.g.~traits,
abundance, range, phylogenetic relatedness, etc.) to build models to
predict interactions between species we haven't observed together before
(Strydom \emph{et al.} 2021). However, the noise of false-negatives
could impact the efficacy of our predictive models and have practical
consequences for answering questions about interactions (de Aguiar
\emph{et al.} 2019). This data constraint is amplified as the
interaction data we have is geographically biased toward the usual
suspects (Poisot \emph{et al.} 2021). We therefore need a statistical
approach to assessing these biases in the observation process and their
consequences for our understanding of interaction networks.

The intrinsic properties of ecological communities create several
challenges for sampling: first, species are not observed with equal
probability---we are much more likely to observe a species of high
abundance than one of very low abundance (Poisot \emph{et al.} 2015).
Canard \emph{et al.} (2012) presents a null model of food-web structure
where species encounter one-another in proportion to each species'
relative-abundance. This assumes that there are no associations in
species co-occurrence due to an interaction (perhaps because this
interaction is ``important'' for both species; Cazelles \emph{et al.}
(2016)), but in this paper we later show increasing strength of
associations leads to increasing probability of false-negatives in
interaction data, and that these positive associations are rampant in
existing network data. Second, observed co-occurrence is often equated
with meaningful interaction strength, but this is not necessarily the
case (Blanchet \emph{et al.} 2020)---a true ``non-interaction'' would
require that neither of two species, regardless of whether they
co-occur, ever exhibit any meaningful effect on the fitness of the
other. So, although co-occurrence is not directly indicative of an
interaction, it \emph{is} a precondition for an interaction.

Here, we illustrate how our confidence that a pair of species never
interacts highly depends on sampling effort, and suggest that surveys of
species interactions can benefit from simulation modeling of the
sampling process. We demonstrate that the realized false-negative rate
of interactions is directly related to the relative abundance of the
species pool, and demonstrate how simulation can be used to produce a
null estimate of the false-negative rate given total sampling effort
(the total count of all individuals of all species seen). We show that
positive associations in co-occurrence data can increase the realized
number of false-negatives, and demonstrate these positive associations
are rampant in network datasets. We conclude by recommending that the
simulation of sampling effort and species occurrence can and should be
used to help design surveys of species interaction diversity (Moore \&
McCarthy 2016), and by advocating use of null models like those
presented here as a tool for both guiding design of surveys of species
interactions and for modeling detection error in predictive models.

@@

We start with a naive model of interaction detection: we assume that
every interacting pair of species is incorrectly observed as
not-interacting with an independent and fixed probability, which we
denote \(p_fn\) and subsequently refer to as the False-Negative Rate
(FNR). If we observe the same species not-interacting \(N\) times, then
the probability of a true-negative (denoted \(p_{tn}\)) is given by
\(p_{tn}=1-(p_{fn})^N\). This relation (the probability-mass-function of
geometric distribution, a special case of the negative-binomial
distribution) is shown in fig.~\ref{fig:geometric}(A) for varying values
of \(p_{fn}\) and illustrates a fundamental link between our ability to
reliably say an interaction doesn't exist---\(p_{tn}\)---and the number
of times \(N\) we have observed a given species. In addition, note that
there is no non-zero \(p_{fn}\) for which we can ever \emph{prove} that
an interaction does not exist---no matter how many observations of
non-interactions \(N\) we have, \(p_{tn}<1\).

From fig.~\ref{fig:geometric}(A) it is clear that the more often we see
two species co-occurring, but \emph{not interacting}, the more likely
the interaction is a true-negative. This has several practical
consequences: first it means negatives taken outside the overlap of the
range of each species aren't informative because co-occurrence was not
possible, and therefore neither was an interaction. Second, we can use
this relation to compute the expected number of total observations
needed to obtain a ``goal'' number of observations of a particular pair
of species (fig.~\ref{fig:geometric}(B)). As an example, if we
hypothesize that \(A\) and \(B\) do not interact, and we want to see
species \(A\) and \(B\) both co-occurring and \emph{not interacting} 10
times to be confident this is a true negative, then we need an expected
1000 observations of all species if the relative abundances of \(A\) and
\(B\) are both \(0.1\).

@@

\begin{figure}
\hypertarget{fig:geometric}{%
\centering
\includegraphics{./figures/fig1.png}
\caption{\textbf{(A)} The probability that an observed interaction is a
true negative (y-axis) given how many times it has been sampled as a
non-interaction (x-axis). Each color reflects a different value of
\(p_{fn}\), the false-negative rate (FNR)---this is effectively the cdf
of the geometric distribution. \textbf{(B)} The expected number of total
observations needed (colors) to observe 10 co-occurrences between a
species with relative abundance \(P(A)\) (x-axis) and a second species
with relative abundance \(P(Y)\). \textbf{(C)}: False negative rate
(y-axis) as a function of total sampling effort (x-axis) and network
size, computed using the method described above. For 500 independent
draws from the niche model (Williams \& Martinez (2000)) at varying
levels of species richness (colors) with connectance drawn according to
the flexible-links model (MacDonald \emph{et al.} (2020)) as described
in the main text. For each draw from the niche model, 200 sets of 1500
observations are simulated, for which each the mean false negative rate
at each observation-step is computed. Means denoted with points, with 1
in the first shade and 2 in the second. \textbf{(D)}: Same as
\textbf{(C)}, except using empirical food webs from Mangal database,
where richness. The outlier on \textbf{(D)} is a 714 species
food-web.}\label{fig:geometric}
}
\end{figure}

@@

We now show that the realized FNR changes drastically with sampling
effort due to the intrinsic variation of the abundance of individuals of
each species within a community. We do this by simulating the process of
observation of species interactions, applied both to 243 empirical food
webs from the Mangal database (Banville \emph{et al.} 2021) and random
food-webs generated using the niche model, a simple generative model of
food-web structure that accounts for allometric scaling (Williams \&
Martinez 2000). Our neutral model of observation assumes each observed
species is drawn in proportion to each species' abundance at that place
and time. The abundance distribution of a community can be
reasonably-well described by a log-normal distribution (Volkov \emph{et
al.} 2003). In addition to the log-normal distribution, we also tested
the case where the abundance distribution is derived from power-law
scaling \(Z^{(log(T_i)-1)}\) where \(T_i\) is the trophic level of
species \(i\) and \(Z\) is a scaling coefficient (Savage \emph{et al.}
2004), which yields the same qualitative behavior. The practical
consequence of abundance distributions spanning many orders of magnitude
of abundance is that observing two ``rare'' species interacting requires
two low probability events: observing two rare species \emph{at the same
time}.

To simulate the process of observation, for an ecological network \(M\)
with \(S\) species, we sample abundances for each species from a
standard-log-normal distribution. For each true interaction in the
adjacency matrix \(M\) (i.e. \(M_{ij}=1\)) we estimate the probability
of observing both species \(i\) and \(j\) at a given place and time by
simulating \(n\) observations of all individuals of any a species, where
the species of the individual observed at the \(\{1,2,\dots,n\}\)-th
observation is drawn from the generated log-normal distribution of
abundances. For each pair of species \((i,j)\), if both \(i\) and \(j\)
are observed within the n-observations, the interaction is tallied as a
true positive if \(M_{ij}=1\). If only one of \(i\) or \(j\) are
observed---but not both---in these \(n\) observations, but \(M_{ij}=1\),
this is counted as a false-negative, and a true-negative otherwise. For
each pair of species \((i,j)\), if both \(i\) and \(j\) are observed
within the n-observations, the interaction is tallied as a true positive
if \(M_{ij}=1\). If only one of \(i\) or \(j\) are observed---but not
both---in these \(n\) observations, but \(M_{ij}=1\), this is counted as
a false-negative, and a true-negative otherwise (\(M_{ij} = 0\)).

@@

where the MAP estimate of \((\mu, \phi)\) applied to Mangal data from
(MacDonald \emph{et al.} 2020) is \((\mu=0.086, \phi=24.3)\). All
simulations were done with 500 independent replicates of unique niche
model networks per unique number of observations \(n\). All analyses
presented here are done in Julia v1.8 (Bezanson \emph{et al.} 2015)
using both EcologicalNetworks.jl v0.5 and Mangal.jl v0.4 (Banville
\emph{et al.} 2021) and are hosted at (GITHUB\_LINK\_TODO). Note that
the empirical data, for the reasons described above, very likely already
contains many false negatives, we'll revisit this issue in the final
section.

From fig.~\ref{fig:geometric}(C) it is evident that the number of
species considered in a study is inseparable from the false-negative
rate in that study, and this effect should be taken into account when
designing samples of ecological networks in the future. We see a similar
qualitative pattern in fig.~\ref{fig:geometric}(D) where the FNR drops
off quickly as a function of observation effort, mediated by total
richness. The practical consequence of the bottom row of
fig.~\ref{fig:geometric} is whether the total number of observations of
all species (the x-axis) for the threshold FNR we deem acceptable (the
y-axis) is feasible. This raises two points: first, empirical data on
interactions are subject to the practical limitations of funding and
human-work hours, and therefore existing data tend to fall on the order
of hundreds or thousands observations of individuals per site. Clear
aggregation of data on sampling effort has proven difficult to find and
a meta-analysis of network data and sampling effort seems both pertinent
and necessary, in addition to the effects of aggregation of interactions
across taxonomic scales (Gauzens \emph{et al.} 2013; Giacomuzzo \&
Jordán 2021). This inherent limitation on in-situ sampling means we
should optimize where we sample across space so that for a given number
of samples, we obtain the maximum information possible. Second, what is
meant by ``acceptable'' FNR? This raises the question: does a shifting
FNR lead to rapid transitions in our ability inference and predictions
about the structure and dynamics of networks, or does it produce a
roughly linear decay in model efficacy? We explore this in the next
section.

We conclude this section by advocating for the use of neutral models
similar to above to generate expectations about the number of
false-negatives in a data set of a given size. This could prove fruitful
both for designing surveys of interactions but also because we may want
to incorporate models of imperfect detection error into predictive
interactions models, as Joseph (2020) does for species occurrence
modeling. Additionally, we emphasize that one must consider the context
for sampling---is the goal to detect a particular species (as in
fig.~\ref{fig:geometric}(C)), or to get a representative sample of
interactions across the species pool? These arguments are
well-considered when sampling individual species (Willott 2001), but
have not yet been adopted for designing samples of communities.

@@

This model above doesn't consider the possibility that there are
positive or negative associations which shift the probability of
observing two species together due to their interaction (Cazelles
\emph{et al.} 2016). However, here we demonstrate that the probability
of observing a false negative can be higher if there is some positive
association in the occurrence of species \(A\) and \(B\). If we denote
the probability that we observe the co-occurrence of two species \(A\)
and \(B\) as \(P(AB)\) and if there is no association between the
marginal probabilities of observing \(A\) and observing \(B\), denoted
\(P(A)\) and \(P(B)\) respectively, then the probability of observing
their co-occurrence is the product of the marginal probabilities for
each species, \(P(AB) = P(A)P(B)\). In the other case where there is
some positive strength of association between observing both \(A\) and
\(B\) because this interaction is ``important'' for each species, then
the probability of observation both \(A\) and \(B\), \(P(AB)\), is
greater than \(P(A)P(B)\) as \(P(A)\) and \(P(B)\) are not independent
and instead are positively correlated, i.e.~\(P(AB)> P(A)P(B)\). In this
case, the probability of observing a single false-negative in our naive
model from fig.~\ref{fig:geometric}(A) is \(p_{fn}= 1-P(AB)\), which due
to the above inequality implies \(p_{fn}>1-P(A)P(B)\). This indicates an
increasingly greater probability of a false negative as the strength of
association gets stronger, \(P(AB) \to P(AB) \gg P(A)P(B)\). However,
this still does not consider variation in species abundance in space and
time (Poisot \emph{et al.} 2015). If positive or negative associations
between species structure variation in the distribution of \(P(AB)\)
across space/time, then the spatial/temporal biases induced by data
collection would further impact the realized false negative rate, as the
probability of false negative would not be constant for each pair of
species across sites.

To test for these positive associations in data we scoured Mangal for
datasets with many spatial or temporal replicates of the same system.
For each dataset, we compute the marginal probability \(P(A)\) of
occurrence of each species \(A\) across all networks in the dataset. For
each pair of interacting species \(A\) and \(B\), we then compute and
compare the probability of co-occurrence if each species occurs
independently, \(P(A)P(B)\), to the empirical joint probability of
co-occurrence, \(P(AB)\). Following our analysis above, if \(P(AB)\) is
greater than \(P(A)P(B)\), then we expect our neutral estimates of the
FNR above to underestimate the realized FNR. In fig.~\ref{fig:mangal},
we see the difference between \(P(AB)\) and \(P(A)P(B)\) for the seven
suitable datasets with enough spatio-temporal replicates and a shared
taxonomic backbone (meaning all individual networks use common species
identifiers) found on Mangal to perform this analysis. Further details
about each dataset are reported in tbl.~\ref{tbl:id}.

In each of these datasets, the joint probability of co-occurrence
\(P(AB)\) is decisively greater than our expectation if species co-occur
in proportion to their relative abundance \(P(A)P(B)\). This suggests
that there may not be as many ``neutrally forbidden links'' (Canard
\emph{et al.} 2012) as we might think, and that the reason we do not
have records of interactions between rare species is probably due to
observation error. This has serious ramifications for the widely
observed property of nestedness seen in bipartite networks (Bascompte \&
Jordano 2007)---perhaps the reason we have lots of observations between
generalists is because they are more abundant, and this is particularly
relevant as we have strong evidence that generalism drives abundance
(Song \emph{et al.} 2022), not vice-versa.

\begin{figure}
\hypertarget{fig:mangal}{%
\centering
\includegraphics{./figures/fig2.png}
\caption{The difference between joint-probability of co-occurrence
(\(P(AB)\)) and expected probability of co-occurrence under independence
(\(P(A)P(B)\)) for interacting species for each dataset. The red-dashed
line indicates 0 (no association). Each histogram represents a density,
meaning the area of the entire curve sums to 1. The continuous density
estimate (computed using local smoothing) is shown in grey. The p-value
on each plot is the result of a one-sided t-test comparing the mean of
each distribution to 0.}\label{fig:mangal}
}
\end{figure}

\hypertarget{tbl:id}{}
\begin{longtable}[]{@{}ccccccccc@{}}
\caption{\label{tbl:id}This table describes the datasets used in the
above analysis (Fig 2). The table reports the type of each dataset, the
total number of networks in each dataset \((N)\), the total species
richness in each dataset \((S)\), the connectance of each metaweb (all
interactions across the entire spatial-temporal extent) \((C)\), the
mean species richness across each local network \(\bar{S}\), the mean
connectance of each local network \(\bar{C}\), the mean
\(\beta\)-diversity among overlapping species across all pairs of
network species (\(\bar{\beta}_{OS}\)), and the mean \(\beta\)-diversity
among all species in the metaweb (\(\bar{\beta}_{WN}\)). Both metrics
are computed using KGL \(\beta\)-diversity (Koleff \emph{et al.}
2003)}\tabularnewline
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endfirsthead
\toprule
Network & Type & \(N\) & \(S\) & \(C\) & \(\bar{S}\) & \(\bar{C}\) &
\(\bar{\beta}_{OS}\) & \(\bar{\beta}_{WN}\)\tabularnewline
\midrule
\endhead
Kopelke \emph{et al.} (2017) & Food Web & 100 & 98 & 0.037 & 7.87 &
0.142 & 1.383 & 1.972\tabularnewline
Thompson \& Townsend (2000) & Food Web & 18 & 566 & 0.014 & 80.67 &
0.049 & 1.617 & 1.594\tabularnewline
Havens (1992) & Food Web & 50 & 188 & 0.065 & 33.58 & 0.099 & 1.468 &
1.881\tabularnewline
Ponisio \emph{et al.} (2017) & Pollinator & 100 & 226 & 0.079 & 23.0 &
0.056 & 1.436 & 1.870\tabularnewline
Hadfield \emph{et al.} (2014) & Host-Parasite & 51 & 327 & 0.085 & 32.71
& 0.337 & 1.477 & 1.952\tabularnewline
Closs \& Lake (1994) & Food Web & 12 & 61 & 0.14 & 29.09 & 0.080 & 1.736
& 1.864\tabularnewline
CaraDonna \emph{et al.} (2017) & Pollinator & 86 & 122 & 0.18 & 21.42 &
0.312 & 1.527 & 1.907\tabularnewline
\bottomrule
\end{longtable}
